---
title: "Time Series in R"
output:
  html_document:
    toc: yes
    df_print: paged
  html_notebook:
    toc: yes
---


### Base R Object: `ts`

Regularly spaced time series data can be represented as objects of class `ts`.
R modeling functions that work with `ts` objects include `arima`, `forecast`, `strucchange`, etc.

Creating a `ts` object:

```{r, eval=FALSE}
y <- ts(data = 1:12, frequency = 12, start=c(1990,1), end=c(1990,12))
```


`ts` objects have their own plot method (`plot.ts`).

```{r} 
# sunspots is a `ts` object
plot(sunspots, main = "Monthly Sunspot Numbers, 1749-1983") 
```

To subset a `ts` object and preserve the date information, use the `window()` function.

```{r}
window(sunspots, start = c(1950,1), end = c(1950,12))
```

Creating lagged data:

```{r, eval=FALSE}
stats::lag(x)  # lag the data by 1 period
stats::lag(x, k = 12)  # lag the data by 12 periods
```

Creating differences:

```{r, eval=FALSE}
base::diff(x)  # first difference
base::diff(x, lag=12) # annual difference
```

To combine multiple `ts` objects into a single object, use the `cbind()` function.
The combined object is assigned the additional class `mts` (multiple time series).

```{r, eval=FALSE}
z <- cbind(x, y)
```

Plot of `mts` objects puts each series in a separate panel. 

```{r}
plot(cbind(sunspots, base::diff((sunspots))), main = "Sunspots")
```


### Ordered Observations: `zoo` 

The `ts` class is quite limited. Irregular spacing data cannot be represented using the `ts` object. 

The `zoo` (abbr. Zeileis' Ordered Observations) class was designed to handle time series data with an arbitrary ordered time index. 
A `zoo` object essentially attaches date information stored in a vector with data information stored in a matrix.

```{r, message=FALSE}
library(zoo)
```

To create a `zoo` object one needs an ordered time index and data. 
Typically, the time index is an object of class `Date`, `POSIXct`, `yearmon`, `yearqtr`, or `timeDate`.

```{r, eval=FALSE}
t <- seq(as.Date("1993/3/1"),  as.Date("2003/3/1"), "months")
x <- cumsum(rnorm(length(t)))
z <- zoo(x = x, order.by = t)
```

Creating `zoo` object from `ts`:

```{r}
sunspots_zoo <- as.zoo(sunspots)
class(sunspots_zoo)
```

The `plot()` function can be used to plot `zoo` objects. 

```{r}
plot(sunspots_zoo, main="Sunspots, 1749-1983")
```

The time index and data can be extracted using the `index()` and `coredata()` functions.

```{r, eval=FALSE}
index(x)
coredata(x)
```


To combine multiple `zoo` objects, use either `cbind()` or `merge()`. 
Use `cbind()` when combining `zoo` objects with the same time index.
Use `merge()` when the objects have different time indexes. 

```{r, eval=FALSE}
merge(x, y) # default outer join
merge(x, y, join="inner")
merge(x, y, join="right", fill=na.locf) # last observation carry forward
```

Dealing with missing values:

```{r, eval=FALSE}
na.locf(xts_object, na.rm = TRUE, fromLast = FALSE, maxgap = Inf) 
na.fill(object, fill, ...)  # replacing NAs
na.trim(object, ...)  # trimming NAs
na.omit(object, ...)  # removing NAs
na.approx(object, ...)  # linear interpolation
```


Reading and writing zoo series:

```{r, eval=FALSE}
read.zoo("foo.csv", FUN = as.yearmon, index.column = 1)
read.zoo("foo.tab", format = "%m/%d/%Y")
```


`tseries` is a time series analysis and computational finance package that works with `zoo` objects. 

```{r, message=FALSE}
library(tseries)
```

Download historical stock quote from Yahoo Finance:

```{r, message=FALSE}
sbux_c <- get.hist.quote(instrument = "MSFT", quote = "Close", retclass = "zoo")
```
```{r}
plot(sbux_c)
```

### Extensible Time Series: `xts`

`xts` (eXtensible Time Series) is a powerful package that provides an extensible time series class, enabling uniform handling of many R time series classes by extending `zoo`.

```{r, message=FALSE}
library(xts)
```

Creating `xts` object:

```{r, eval=FALSE}
y <- xts(x = 1:100, order.by = Sys.Date() + 1:100)
```

Converting to `xts`:

```{r sunspots}
sunspots_xts <- as.xts(sunspots)
plot(sunspots_xts)
```

`xts` object also works with `ggplot2`.

```{r}
library(ggplot2)
autoplot(sunspots_xts)
```

Subsetting:

```{r eval=FALSE}
sunspots_xts["1900"]
sunspots_xts["1900/1901-07"]
sunspots_xts["1900-03/1901-07"]
```

Periodicity:

```{r eval=FALSE}
periodicity(sunspots_xts)
to.yearly(sunspots_xts)  # convert to yearly OHLC
to.period(sunspots_xts, period = "years") 
```

Aggregating:

```{r}
sunspots_xts_yr <- apply.yearly(sunspots_xts, FUN = mean)
plot(sunspots_xts_yr)
```

Continuous rolling windows:

```{r}
sunspots_xts_ma <- rollapply(sunspots_xts, width=120, FUN=mean)
plot(merge(sunspots_xts, sunspots_xts_ma))
```

### Tidy Time Series: `tsibble`

`tsibble` is a variation of tidy data object `tibble` that deals with temporal data. 

```{r}
library(fpp3)  # a short cut to load all tidy-related packages
```

Creating `tsibble` object:

```{r, eval=FALSE}
y <- tsibble(
  Year = 2015:2019,
  Observation = c(123, 39, 78, 52, 110),
  index = Year
)
```

Converting to `tsibble` object:

```{r, eval=FALSE}
z %>%
  mutate(Month = yearmonth(Month)) %>%
  as_tsibble(index = Month)
```

Generating white-noise:

```{r}
tsibble(period = 1:100, wn = rnorm(100), index = period) %>% autoplot(wn)
```

Aggregating over calendar periods:

```{r}
vic_elec %>%
  index_by(date = ~ as_date(.)) %>% 
  summarise(
    ttl_demand = sum(Demand, na.rm = T),
    avg_temp = mean(Temperature, na.rm = T)
    )
```

Tidy time series plot:

```{r, ansett}
ansett %>% 
  filter(Airports == "MEL-SYD") %>%
  autoplot(Passengers)
```

Seasonal plots:

```{r, PBS}
PBS %>%
  filter(ATC2 == "A10") %>%
  select(Month, Concession, Type, Cost) %>%
  summarise(TotalC = sum(Cost)) %>%
  mutate(Cost = TotalC / 1e6) -> a10

a10 %>% gg_season(Cost) 
```

Seasonal subseries plots:

```{r}
a10 %>% gg_subseries(Cost)
```

Correlation matrix: 

```{r tourism, message=FALSE}
library(GGally)
tourism %>%
  group_by(State) %>%
  summarise(Trips = sum(Trips)) %>%
  pivot_wider(values_from=Trips, names_from=State) %>%
  GGally::ggpairs(columns = 2:9)
```

Lag plots:

```{r}
aus_production %>%
  filter(year(Quarter) >= 2000) %>% 
  gg_lag(Beer, geom = "point")
```

Autocorrelation and partial-autocorrelation:

```{r}
aus_production %>% feasts::ACF(Beer) %>% autoplot()
aus_production %>% feasts::PACF(Beer) %>% autoplot()
```








